---
title: "Learning to Float? Currency Management in Emerging Markets Under the Global Financial Cycles"
subtitle: "Revisiting Calvo and Reinhart (2002) through the 2010–2018 Era"
author:
  - name: |
      Tony De Amorim \tabularnewline
      \small Paris-Dauphine University, PSL \tabularnewline
      \small \href{mailto:tony.de-amorim@dauphine.eu}{Contact} -- \href{https://github.com/TonyDeAmorim}{GitHub} -- \href{https://www.linkedin.com/in/tony-de-amorim-756818257}{LinkedIn}

date: last-modified
date-format: "MMMM YYYY"
abstract: |
  This essay investigates the de facto exchange rate rigidity in Emerging Market Economies (EMEs) during the post-2008 era (2010–2018). Revisiting the framework of Calvo and Reinhart (2002), we analyze currency management under the influence of the Global Financial Cycle, deliberately selecting a pre-pandemic window to isolate market dynamics from recent structural noise. Rather than relying on a static historical baseline, we calibrate a relative variability threshold of 2.13% based on the actual behavior of benchmark pure floaters during this specific decade. The findings illustrate an illusion of free floating: several EMEs continue to exhibit stability patterns that diverge significantly from true floaters, sometimes acting as "Rigid Anchors" to mitigate imported volatility. While not structured as a formal working paper, this essay employs graphical representations and concise analytical frameworks to render complex macroeconomic concepts accessible to a broader audience.
keywords: [Fear of Floating, Exchange Rate Regimes, Global Financial Cycle, Emerging Markets]
format:
  pdf:
    geometry:
      - left=25mm
      - right=25mm
      - top=30mm
      - bottom=30mm
    papersize: a4
    number-sections: true
    fig-pos: 'H'
    fontsize: 11pt
    interlinespacing: 1.2
    indent: true
bibliography: Refferences.bib
csl: apa.csl
---


```{r}
#| label: library import
#| echo: false
#| message: false
#| warning: false

library(ggplot2)
library(ggthemes)
library(tidyverse)
library(readr)
library(yfR)
library(wbstats)
library(priceR)
library(patchwork)
library(olsrr)
library(yfR)
library(zoo)
library(knitr)
library(kableExtra)
library(readr)
library(gt)
library(countrycode)
library(readxl)
```

```{r}
#| label: Import for devise function
#| echo: false
#| message: false
#| warning: false


Import_DEV <- function(DEV) {
  Ticker <- str_c(DEV, "=X")

  df <- yf_get(
    tickers = Ticker,
    first_date = "2010-01-01",
    last_date = "2018-01-01",
    thresh_bad_data = 0
  )

  df <- df |>
    select(ref_date, price_close) |>
    mutate(month = floor_date(ref_date, "month")) |>
    group_by(month) |>
    slice_tail(n = 1) |>
    ungroup() |>
    mutate(
      diff_price = ((price_close - lag(price_close)) / lag(price_close)) * 100,
      country = str_c("USD_", DEV)
    ) |>
    filter(!is.na(diff_price)) |>
    select(month, country, diff_price)

  return(df)
}
```

```{r}
#| label: Var plot dev function
#| echo: false
#| message: false
#| warning: false

Var_Plot_DEV <- function(df, name, tr) {
  ggplot(df, aes(x = month, y = .data[[name]])) +
    geom_line() +
    geom_hline(yintercept = tr, linetype = "dashed", color = "red", alpha = 0.6) +
    geom_hline(yintercept = -tr, linetype = "dashed", color = "red", alpha = 0.6) +
    theme_bw(base_size = 8) +
    labs(
      title =  name,
      y = "Variation (%)",
      x = "Date"
    )
}
```

*Note : Throughout this essay, we will use the formulation used in @ahmed2022monetary, referring to developing countries as Emerging Market Economies (EMEs), as this wording is useful in both its simplicity and clarity.*

# Introduction


Consistent with the foundational work of @calvo2002fear, this short essay investigates the de facto exchange rate rigidity of Emerging Market Economies (EMEs) during the post-2008 era. Specifically, it analyzes the propensity of these nations to smooth their currency fluctuations to mitigate imported macroeconomic shocks, particularly those driven by the Global Financial Cycle. As demonstrated in recent literature, notably by @miranda2020us, the Global Financial Cycle acts as a single global factor—closely tied to U.S. monetary policy and global risk aversion—that drives international credit conditions and capital flows, significantly constraining the monetary autonomy of EMEs regardless of their official exchange rate regimes. Although this is not an official working paper, this essay uses visual representations and a concise analytical framework to present this complex macroeconomic phenomenon on an understandable basis.

In the original literature, *Fear of Floating* is defined as a phenomenon where EMEs, despite officially adopting flexible exchange rate regimes, remain highly sensitive to currency fluctuations due to concerns regarding dollar-denominated debt and inflationary pressures. Because the foundational research by @calvo2002fear predates the structural shifts that followed the 2008 crisis, the persistence of such an "illusion of free floating" under the intense pressures of the Global Financial Cycle remains a critical empirical question. By benchmarking a panel of EMEs against major free-floating currencies during the 2010-18 period, this study replicates and extends the original methodology to compare declared exchange regimes against realized market volatility.

The structure of this essay is as follows: Section 2 details the identification of exchange rate regimes and the construction of the sample. Section 3 explains the methodology used to endogenize a modern variation threshold of 2.13%. Section 4 applies this framework to EMEs to evaluate the persistence of Fear of Floating relative to historical benchmarks. Finally, Section 5 contextualizes these findings within modern theoretical frameworks, such as the Global Financial Cycle, and offers concluding remarks.

# Methodology and Sample Selection

This section details the construction of the study sample, which was developed by cross-referencing the foundational country list from @calvo2002fear with the IMF’s Annual Report on Exchange Arrangements and Exchange Restrictions (@imf2024areaer) database. Rather than attempting a broad contemporary analysis that would be skewed by the unprecedented structural noise of the COVID-19 pandemic, we deliberately focus on the 2010–2018 window. This specific period serves as an ideal macroeconomic laboratory: it captures the post-2008 Global Financial Cycle at its peak, allowing us to observe the de facto exchange rate rigidity of emerging markets under a stabilized global environment, prior to the extreme volatility of the 2020s. To ensure robust observations within this era, we selected economies that maintained a stable classification of either "floating" or "free floating" during this timeframe. These precise exchange rate regimes are officially defined as such by the IMF:

- **Floating** : Exchange rates are largely determined by the market, with no verifiable or predictable trend. Intervention in the foreign exchange market can be direct or indirect, and serves to moderate the pace of changes and prevent excessive fluctuations in the exchange rate. They may exhibit varying degrees of volatility, depending on the magnitude of the shocks affecting the economy.
- **Free Floating** : Interventions are carried out only in exceptional cases to address market disruptions, and provided that the authorities have demonstrated that such interventions were limited to no more than three instances over the past six months, with each lasting no longer than three business days. 

This dual-category approach is essential for distinguishing advanced economies, characterized by minimal intervention, from emerging markets, which may exhibit controlled volatility. The latter are prominently represented by emerging market economies (EMEs) such as Brazil, India, and Peru. While these countries officially allow the market to determine the exchange rate, they nevertheless reserve the right to intervene to mitigate currency fluctuations. It is therefore within this group that we expect to observe the clearest evidence of de facto exchange rate rigidity and the illusion of free floating. The officially declared exchange regime for each country during our target window is reported in @tbl-Exchange-Rates-Regime. 

```{r}
#| label: tbl-Exchange-Rates-Regime
#| tbl-cap: "Classification of exchange rate regimes"
#| echo: false
#| message: false
#| warning: false


exchange_regime_raw <- read_delim("Ressources/exchange_regime.csv", 
                                  delim = ";", 
                                  escape_double = FALSE, 
                                  trim_ws = TRUE)

free_floaters <- exchange_regime_raw |>
  select(Free_Floating = 1) |>
  filter(!is.na(Free_Floating))

floaters <- exchange_regime_raw |>
  select(Floating = 3) |>
  filter(!is.na(Floating))

n_free <- nrow(free_floaters)
n_float <- nrow(floaters)
max_rows <- max(n_free, n_float)

if (n_free < max_rows) {
  free_floaters <- bind_rows(free_floaters, 
                             tibble(Free_Floating = rep("", max_rows - n_free)))
}

if (n_float < max_rows) {
  floaters <- bind_rows(floaters, 
                        tibble(Floating = rep("", max_rows - n_float)))
}

table_display <- bind_cols(free_floaters, floaters)

table_display |>
  kable(
    format = "latex",
    booktabs = TRUE,
    col.names = c("Free Floating", "Floating"),
    align = "l"
  ) |>
  kable_styling(
    latex_options = c("hold_position"),
    font_size = 10
  ) |>
  column_spec(1, width = "6cm") |>
  column_spec(2, width = "6cm")


```


# Endogenizing the Variability Threshold

```{r}
#| label: Importing Benchmark data
#| echo: false
#| message: false
#| warning: false

Benchmark <- c("JPY", "EUR", "GBP", "AUD")

Benchmark_df <- map(Benchmark, Import_DEV) |>
  list_rbind() |>
  pivot_wider(names_from = country, values_from = diff_price)
```

```{r}
#| label: Computing new threshold
#| echo: false
#| message: false
#| warning: false

target_probs <- c(0.55, 0.60, 0.65)

threshold_comparison <- map_dfr(target_probs, function(p) {
  tr_value <- Benchmark_df |>
    pivot_longer(cols = starts_with("USD_"), names_to = "currency", values_to = "variation") |>
    group_by(currency) |>
    summarise(indiv_tr = quantile(abs(variation), probs = p, na.rm = TRUE)) |>
    summarise(mean_tr = mean(indiv_tr)) |>
    pull(mean_tr)
  tibble(target_percent = p * 100, threshold_obtained = round(tr_value, 2))
})

threshold_comparison <- threshold_comparison |>
  add_row(target_percent = 60, threshold_obtained = 2.50) |>
  mutate(label = ifelse(threshold_obtained == 2.50, "Original (2002)", "Modern (Endogenous)"))

```


In their seminal work, Calvo and Reinhart (2002) utilized a fixed 2.5% threshold to evaluate the exchange rate stability of benchmark economies (specifically the G3 and Australia) relative to EMEs. By calculating the frequency of monthly fluctuations within this band, they identified significant disparities. Their findings revealed that many self-declared floaters exhibited remarkably low exchange rate variance. However, given the structural shifts in global financial markets over the last two decades, a static 2.5% band may no longer accurately define the modern standard for currency flexibility. Consequently, this essay adopts an empirical strategy to endogenize the threshold based on contemporary data. Taking the historical findings as a baseline, where benchmark floaters exhibited a 60% probability of remaining within a 2.5% band, we reverse engineer this logic to capture the "normal" market volatility and inherent noise of modern benchmark currencies. Rather than merely reproducing a static historical figure, we determine the specific threshold that replicates this 60% probability for each of our modern benchmark group, effectively defining the standard variability for a free-floating regime, and then take the average to establish a new benchmark to which the EMEs in our sample can be compared. We thus seek to identify a new threshold $x$ such that:
$$
P(|\Delta e _{it}|<x)= 0.60
$$ {#eq-Prob}

Where $e_{it}$ represents the exchange rate of country $i$ against the USD at time $t$. We apply this method to two of the original currencies, the Japanese Yen (JPY) and the Australian Dollar (AUD). We substituted the German Deutsche Mark with the Euro (EUR) and incorporated the British Pound (GBP), which is widely considered a robust free-floating currency. Daily data from Yahoo Finance were aggregated into monthly iterations using the end-of-period value. The use of end-of-period values is preferred over monthly averages to ensure that the exchange rate reflects a market spot price at a specific point in time, rather than a smoothed aggregate. This choice is methodologically consistent with the framework used in Calvo and Reinhart (2002), which evaluates the variability of "stock" variables. By using EOP data, the relative variations $\Delta e_{it}$ in @eq-Variations represent the actual price movement between two discrete points, capturing the full extent of market shocks that occur at the end of the month which might otherwise be masked by the arithmetic mean of a period average. To compute the variations, we follow the framework of the original work by using the absolute value of monthly changes:
$$
\Delta e_{it} = \frac{e_{it} - e_{it-1}}{e_{it-1}} \cdot 100
$$ {#eq-Variations}

This metric captures the discrete relative variation of the exchange rate, serving as a proxy for financial stress. Higher percentage changes reflect increased exchange rate variability. While volatility is statistically measured as the dispersion of returns over time, $\Delta e_{{it}}$ serves as a direct proxy for financial stress. By focusing on the magnitude of these discrete movements, we can identify episodes where the exchange rate deviates significantly from its previous level, signaling potential market instability. Furthermore, by focusing on relative rather than absolute changes, this measure ensures comparability across different currency pairs regardless of their nominal price levels. After using our method and using the 60% probability as a first estimate, we obtain a 2.13% threshold and represented it for our benchmark on @fig-Benchmark-variation.

```{r}
#| label: fig-Benchmark-variation
#| fig-cap: "Benchmark Currencies Volatility Analysis"
#| echo: false
#| message: false
#| warning: false



currencies <- c("USD_JPY", "USD_EUR", "USD_GBP", "USD_AUD")

Bch_plots <- lapply(currencies, function(x) Var_Plot_DEV(Benchmark_df, x, 2.13))

wrap_plots(Bch_plots, ncol = 2) +
  plot_annotation(
    caption = "Threshold of 2.13% derived from the 60% quantile of observations
    Source: Yahoo Finance data aggregated by author",
    theme = theme(
      plot.caption = element_text(size = 6)
    )
  )
```

# Empirical Analysis: Measuring Fear of Floating in EMEs

```{r}
#| label: Importing EMEs free floaters data
#| echo: false
#| message: false
#| warning: false

EMEs_FF <- c("CAD", "CLP", "IDR", "MXN", "NOK", "SEK")

EMEs_FF_df <- map(EMEs_FF, Import_DEV) |>
  list_rbind() |>
  pivot_wider(names_from = country, values_from = diff_price)
```

```{r}
#| label: Importing EMEs floaters data
#| echo: false
#| message: false
#| warning: false

EMEs_FL <- c("BRL", "COP", "INR", "ILS", "KES", "MYR", "NZD", "PEN", "PHP", "ZAR", "KRW", "UGX", "UYU")

EMEs_FL_df <- map(EMEs_FL, Import_DEV) |>
  list_rbind() |>
  pivot_wider(names_from = country, values_from = diff_price)
```

```{r}
#| label: Computing probabilities for free floaters
#| echo: false
#| message: false
#| warning: false


thresholds <-
  threshold_comparison |>
  pull(threshold_obtained)

names(thresholds) <- paste0("tr_", threshold_comparison$target_percent)

FF_stability_multi <- EMEs_FF_df |>
  pivot_longer(
    cols = starts_with("USD_"),
    names_to = "country",
    values_to = "variation"
  ) |>
  mutate(
    p_55 = abs(variation) < thresholds["tr_55"],
    p_60 = abs(variation) < thresholds["tr_60"],
    p_65 = abs(variation) < thresholds["tr_65"],
    p_ORG = abs(variation) < 2.5
  ) |>
  group_by(country) |>
  summarise(
    across(starts_with("p_"), ~ round(mean(.x, na.rm = TRUE) * 100, 2))
  ) |>
  arrange(desc(p_60))
```

```{r}
#| label: Computing probabilities for floaters
#| echo: false
#| message: false
#| warning: false


FL_stability_multi <- EMEs_FL_df |>
  pivot_longer(
    cols = starts_with("USD_"),
    names_to = "country",
    values_to = "variation"
  ) |>
  mutate(
    p_55 = abs(variation) < thresholds["tr_55"],
    p_60 = abs(variation) < thresholds["tr_60"],
    p_65 = abs(variation) < thresholds["tr_65"],
    p_ORG = abs(variation) < 2.5
  ) |>
  group_by(country) |>
  summarise(
    across(starts_with("p_"), ~ round(mean(.x, na.rm = TRUE) * 100, 2))
  ) |>
  arrange(desc(p_60))
```

```{r}
#| label: Computing probabilities for benchmark
#| echo: false
#| message: false
#| warning: false


Benchmark_stability_multi <- Benchmark_df |>
  pivot_longer(
    cols = starts_with("USD_"),
    names_to = "country",
    values_to = "variation"
  ) |>
  mutate(
    p_55 = abs(variation) < thresholds["tr_55"],
    p_60 = abs(variation) < thresholds["tr_60"],
    p_65 = abs(variation) < thresholds["tr_65"],
    p_ORG = abs(variation) < 2.5
  ) |>
  group_by(country) |>
  summarise(
    across(starts_with("p_"), ~ round(mean(.x, na.rm = TRUE) * 100, 2))
  ) |>
  arrange(desc(p_60))
```

We apply our baseline threshold of 2.13% to the Emerging Market Economies (EMEs). The graphical representations for both floaters and free floaters, provided in the [Appendix](#sec-one), reveal highly heterogeneous patterns of variation. The core objective of this analysis is to compute the probability of each EME remaining within this 2.13% band to allow for a direct comparison with the original findings of Calvo and Reinhart (2002).The selection of the 55% and 65% quantiles provides a symmetrical ±5% interval around our primary 60% baseline, allowing for a rigorous sensitivity test of the results. This approach establishes a variability gradient that captures the persistence of stability patterns across varying degrees of strictness. While the 55% quantile (1.80%) imposes a highly restrictive definition to test the most intense instances of currency management, the 65% quantile (2.33%) ensures the findings remain consistent as the threshold approaches the historical 2.50% mark. While the persistent divergence between Benchmarks and Floaters across these intervals suggests a deliberate stabilization effort, this statistical alignment remains a proxy for policy intent rather than a formal confirmation of active intervention, as it does not explicitly control for foreign exchange reserve movements or interest rate adjustments. Finally, we include the original 2.5% band as a historical reference. Our results are detailed in @tbl-Probability-calcul. 

```{r}
#| label: tbl-Probability-calcul
#| tbl-cap: "Volatility of Exchange Rates in Different Regimes"
#| echo: false
#| message: false
#| warning: false

custom_names <- c(
  "GBP" = "United Kingdom",
  "JPY" = "Japan",
  "EUR" = "Euro Area",
  "AUD" = "Australia",
  "IDR" = "Indonesia",
  "CAD" = "Canada",
  "MXN" = "Mexico",
  "SEK" = "Sweden",
  "CLP" = "Chile",
  "NOK" = "Norway",
  "KES" = "Kenya",
  "PHP" = "Philippines",
  "PEN" = "Peru",
  "UGX" = "Uganda",
  "INR" = "India",
  "UYU" = "Uruguay",
  "ILS" = "Israel",
  "MYR" = "Malaysia",
  "KRW" = "South Korea",
  "COP" = "Colombia",
  "NZD" = "New Zealand",
  "BRL" = "Brazil",
  "ZAR" = "South Africa"
)

all_results <- bind_rows(
  Benchmark_stability_multi |> mutate(group = "Benchmark"),
  FF_stability_multi |> mutate(group = "Free Floaters"),
  FL_stability_multi |> mutate(group = "Floaters")
) |>
  mutate(
    curr_code = str_extract(country, "[A-Z]{3}$"),
    country_name = recode(curr_code, !!!custom_names)
  ) |>
  arrange(group, desc(p_60))


all_results <- all_results |>
  mutate(group = factor(group, levels = c("Benchmark", "Floaters", "Free Floaters"))) |>
  arrange(group, country)

group_counts <- all_results |> count(group)
tab_final <- all_results |>
  select(Country = country_name, p_55, p_60, p_65, p_ORG) |>
  mutate(across(starts_with("p_"), ~ sprintf("%.1f", .)))

tab_final |>
  kable(
    format = "latex",
    booktabs = TRUE,
    col.names = c("Country", "Within ±1.80% band", "Within ±2.13% band", 
                  "Within ±2.33% band", "Within ±2.50% band"),
    align = "lcccc"
  ) |>
  kable_styling(
    latex_options = c("H"),
    font_size = 10
  ) |>
  pack_rows("Benchmark", 1, group_counts$n[1], indent = FALSE, bold = TRUE) |>
  pack_rows("Floaters", group_counts$n[1] + 1, group_counts$n[1] + group_counts$n[2], 
            indent = FALSE, bold = TRUE) |>
  pack_rows("Free Floaters", sum(group_counts$n[1:2]) + 1, nrow(tab_final), 
            indent = FALSE, bold = TRUE) |>
  add_header_above(c(" " = 1, "Probability that the monthly change is:" = 4), bold = TRUE) |>
  footnote(
  general = "Note: Data 2010-2018. Thresholds are endogenously determined. 2.50% is the Calvo-Reinhart (2002) reference.",
  general_title = "",
  footnote_as_chunk = TRUE,
  threeparttable = TRUE
)
```

Our findings reveal a notable divergence between official exchange rate mandates and realized market volatility. Certain nations classified as "Free Floaters," most notably Indonesia, exhibit stability levels that appear to deviate from the Benchmark group, which may suggest that forms of de facto management persist even under officially flexible frameworks. The relative positioning of Emerging Market Economies remains largely consistent across these ours 1.80% and 2.33% intervals, highlighting two distinct volatility profiles within the sample.

The "Rigid Anchor" profile, represented by nations such as Kenya and the Philippines, maintains stability levels that surpass even the most stable advanced economy Benchmarks. These countries display probabilities of remaining within the exchange rate band that often exceed 75% even at the most restrictive 1.80% threshold, hinting at a potential policy preference for smoothing minor currency fluctuations. Conversely, the "Progressive Flexible" profile, observed in economies like Mexico and Chile, demonstrates a volatility profile that aligns more closely with the behavior of pure floaters. In these cases, stability levels decrease as the bands tighten, which might indicate that these EMEs have integrated exchange rate variability as a functional tool within their broader economic framework.

This consistency across varying thresholds suggests that our observations likely capture underlying policy smoothing or structural management rather than merely reflecting a specific choice of exogenous parameters. The persistence of these patterns across different quantiles points toward the possibility that currency management is influenced by structural vulnerabilities, such as debt-dollarization or the necessity to mitigate the effects of the Global Financial Cycle. While these rankings hold regardless of the specific width of the variation band, these results should be interpreted as a preliminary framework, as the discrete variation of the exchange rate serves as a proxy and does not explicitly disentangle market-driven movements from active policy intervention.

# Revisiting the Trilemma through the Global Financial Cycle

Our empirical results must be interpreted within the evolving framework of international macroeconomics. While the classical Impossible Trinity of @mundell1963capital suggests that flexible exchange rates grant monetary independence under free capital movement, recent literature on the Global Financial Cycle challenges this view. @rey2015dilemma argues that the international system faces a dilemma rather than a trilemma, as exchange rate flexibility fails to provide autonomy when global credit conditions and capital flows are driven by a single global factor. Closely tied to the US monetary policy stance and global risk aversion, this factor creates a significant constraint for EMEs. As @miranda2020us demonstrated, the US dollar acts as the primary engine for this cycle, exposing even floating regimes to imported volatility regardless of their domestic fundamentals. 

This theoretical framework contextualizes the results in @tbl-Comp-calcul, which compares contemporary stability probabilities at the 2.5% threshold with the original findings of @calvo2002fear. The observed decrease in stability across most groups suggests that the modern financial environment may be more volatile than in the early 2000s. However, these results should be interpreted with caution due to the limited sample size in certain categories, such as the 'Become FF' group (N=2). While every category exhibits increased variability, the shift appears most pronounced for currencies that transitioned to free-floating regimes. Conversely, countries moving toward managed arrangements show more resilient stability levels, which may reflect policy efforts to dampen the fluctuations driven by the global financial cycle.

```{r}
#| label: tbl-Comp-calcul
#| tbl-cap: "Stability Variations by Regime Transition (Threshold: 2.5%)"
#| echo: false
#| message: false
#| warning: false
Comparison_table <- read_excel("Ressources/Comparison_table.xlsx")

transition_summary <- Comparison_table |>
  filter(!Country %in% c("Australia", "Japan")) |>
  mutate(
    Exchange_rate_2002 = as.numeric(Exchange_rate_2002),
    Exchange_rate_2010_19 = as.numeric(Exchange_rate_2010_19)
  ) |>
  mutate(
    Transition = case_when(
      Regime_2002 == "FF" & Regime_2010_19 == "FF" ~ "Stay FF",
      Regime_2002 == "FL" & Regime_2010_19 == "FL" ~ "Stay FL",
      Regime_2002 == "FL" & Regime_2010_19 == "FF" ~ "Become FF",
      Regime_2002 == "FF" & Regime_2010_19 == "FL" ~ "Become FL"
    ),
    Variation = Exchange_rate_2010_19 - Exchange_rate_2002
  ) |>
  group_by(Transition) |>
  summarise(
    n = n(),
    Countries = paste(Country, collapse = ", "),
    Mean_2002 = mean(Exchange_rate_2002, na.rm = TRUE),
    Mean_2010_19 = mean(Exchange_rate_2010_19, na.rm = TRUE),
    Avg_Change = mean(Variation, na.rm = TRUE)
  ) |>
  arrange(Avg_Change)

transition_summary |>
  kable(
    format = "latex",
    booktabs = TRUE,
    col.names = c("Transition", "N", "Countries", "Mean 2002", "Mean 2010-19", "Change"),
    digits = 1
  ) |>
  kable_styling(latex_options = c("hold_position", "scale_down")) |>
  column_spec(3, width = "6cm") |>
  column_spec(6, bold = TRUE) |>
  footnote(
    general = "Note: FF and FL denotes 'Free Floating' and 'Floating' respectively, 'Become FF' and 'Become FL' highlights countries moving that switched between 'Free Floating' and 'Floating' regimes (and inversly) between the two studies.",
    general_title = "",
    threeparttable = TRUE,
    fixed_small_size = 2.5
  )

```

The limitations of a fixed 2.5% threshold are particularly evident when considering the heightened sensitivity of EMEs to external shocks. As @arora2001how demonstrated, changes in U.S. monetary policy transmit directly to emerging markets through sovereign spreads, creating a financial vulnerability that persists regardless of the official exchange rate regime. Given that dependency on dollar-denominated debt remains pervasive, the baseline volatility of the global economy has shifted significantly since the early 2000s. Consequently, relying on a static 2.5% benchmark fails to account for these structural changes and the increased sophistication of modern currency management. By endogenizing our thresholds at 2.13%, we move away from an arbitrary historical figure to define *Fear of Floating* relative to contemporary market conditions. This approach ensures that our results capture the actual policy intent of central banks rather than simply measuring their distance from an outdated international norm. 

These tighter bands provide a more nuanced detection of currency management, as shown in @tbl-Probability-calcul. A striking divergence emerges: while benchmark currencies see their stability probabilities drop sharply as the threshold narrows, a subset of floaters, most notably Kenya and the Philippines, maintain remarkably high stability. This suggests that these central banks are not merely preventing extreme devaluations but are actively smoothing even minor fluctuations. By defining *Fear of Floating* relative to the actual behavior of pure floaters in the contemporary era, we highlight that the intensity of currency management in certain EMEs is significantly higher than a static historical reference would suggest. 

# Concluding Remarks

The findings of this study suggest that *Fear of Floating* may still be a significant factor in the modern international monetary system, though its drivers appear increasingly complex. By endogenizing the variation threshold to 2.13%, this analysis reveals that while global volatility has increased, several EMEs continue to maintain stability levels that diverge from the behavior of pure floaters. The results identify a notable distinction between "Rigid Anchors," such as Kenya and the Philippines, and "Progressive Flexibles," such as Mexico and Chile. These patterns invite further investigation into the shift from a traditional trilemma to a monetary dilemma, where the Global Financial Cycle and US dollar dominance potentially constrain domestic autonomy regardless of the official regime. Ultimately, this evidence suggests that contemporary currency management is influenced less by official labels and more by structural vulnerabilities and the perceived necessity to mitigate imported shocks.

However, these results should be interpreted as a preliminary framework rather than a definitive classification, as several factors may bias the observations. The reliance on the 2010 to 2018 window ensures data consistency but omits the unprecedented volatility of the post-pandemic era. Furthermore, while the discrete variation of the exchange rate serves as a useful proxy, it does not explicitly disentangle market-driven movements from active policy intervention. The omission of foreign exchange reserve data, interest rate volatility and the proportion of dollarized-debt for EMEs remains a significant limitation, as these instruments are critical components of the EME policy toolkit. These constraints invite an extended agenda that incorporates central bank interventions to better differentiate intentional policy smoothing from the passive consequences of global financial conditions.


\newpage

# Appendix : Variations Plots {#sec-one -}

```{r}
#| label: fig-EMEs_FL-variation
#| fig-cap: "EMEs Floaters Currencies Volatility Analysis"
#| echo: false
#| message: false
#| warning: false
#| fig-width: 12
#| fig-height: 9
#| out-width: "90%"

floaters <- c("USD_BRL", "USD_COP", "USD_INR", "USD_ILS", "USD_KES", "USD_KRW", "USD_MYR", "USD_NZD", "USD_PEN", "USD_PHP", "USD_ZAR", "USD_UGX", "USD_UYU")

FL_plots <- lapply(floaters, function(x) Var_Plot_DEV(EMEs_FL_df, x, 2.13))

wrap_plots(FL_plots, ncol = 4) +
  plot_layout(guides = "collect") +
  plot_annotation(
    caption = "Threshold of 2.13% derived from the 60% quantile of observations\nSource: Yahoo Finance data aggregated by author",
    theme = theme(plot.caption = element_text(size = 10))
  ) &
  theme(axis.title = element_blank())
```
 
```{r}
#| label: fig-EMEs_FF-variation
#| fig-cap: "EMEs Free Floaters Currencies Volatility Analysis"
#| echo: false
#| message: false
#| warning: false
#| fig-width: 10
#| fig-height: 8
#| out-width: "70%"


free_floaters <- c("USD_CAD", "USD_CLP", "USD_IDR", "USD_MXN", "USD_NOK", "USD_SEK")

FF_plots <- lapply(free_floaters, function(x) Var_Plot_DEV(EMEs_FF_df, x, 2.13))

wrap_plots(FF_plots, ncol = 2) +
  plot_annotation(
    caption = "Threshold of 2.13% derived from the 60% quantile of observations
    Source: Yahoo Finance data aggregated by author",
    theme = theme(
      plot.caption = element_text(size = 10)
    )
  )
```

# Bibliography